A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-17 paper/draft.tex
no way of knowing data from party members
plot summary partisanship

\subsection{Locking-In Wealthy Patrons}

@@

The socioeconomic status of voters is not, however, the only consideration that politicians take into place. Patronage can be used by politicians to reward political loyalists for resources they have contributed for an electoral campaign, a practice generally denoted as \emph{prebendism}\citep{van2007meet}. In Brazil, studies have found that campaign contributors are more likely to be employed in the local bureaucracy, in the magnitude of a 30 percent increase in the probability of being hired \citep{colonnelli2018patronage}. Additionally, there is strong evidence that employment in the bureaucracy is used as a spoil to reward party loyalists, again with evidence from Brazilian local bureaucracies \citep{brollo2017victor}.

The timing of clientelistic transfers ie key. There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office. In particular, a politician cannot offer job to a voter prior to holding office. Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters. There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.

@@

This is in fact what the data suggests. First, on average party members who enter the bureaucracy are wealthier than their non-partisan counterparts, in particular at the moment that the mayor assumes office. With the RAIS data, it is possible to identify for each and every employee the last wage in the non-municipal sector prior to public sector employment. Figure \ref{fig:wage_differential} plots the median of the last non-municipal sector wage by partisanship status. The gap between partisan and non-partisan wages is the highest for the beginning of the mayoral term, which is often associated with more intense patronage activity.

@@

Additionally, these same workers tend to have a longer experience working in the formal sector prior to employment. Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts. They have experienced both longer bouts of employment in the formal sector, as well as higher compensations in their tenure. While it is clear that these hires are a reward for previous contributions, it is not clear whether a party member who is hired into the bureaucracy is more likely to donate to their party. \emph{How can I show that party members who are hired into the bureaucracy are more likely to donate?}

@@

In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. This visualization includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}

>>>>

\subsection{Wealthy Patrons}

@@

The relative cost of voters is not, however, the only consideration that politicians take into place. Patronage can be used by politicians to reward political loyalists for resources they have contributed for an electoral campaign, a practice generally denoted as \emph{prebendism}\citep{van2007meet}. In Brazil, studies have found that campaign contributors are more likely to be employed in the local bureaucracy, in the magnitude of a 30 percent increase in the probability of being hired \citep{colonnelli2018patronage}. Additionally, there is strong evidence that employment in the bureaucracy is used as a spoil to reward party loyalists, again with evidence from Brazilian local bureaucracies \citep{brollo2017victor}.

The timing of clientelistic transfers is key. There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office. In particular, a politician cannot offer job to a voter prior to holding office. Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters. There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.

@@

While proving the second point is beyond the scope of this paper, there is strong evidence that the first dynamic, that patronage disproportionately targets wealthy patrons, is true. First, on average party members who enter the bureaucracy are wealthier than their non-partisan counterparts, in particular at the moment that the mayor assumes office. With the RAIS data, it is possible to identify for each and every employee the last wage in the non-municipal sector prior to public sector employment. Figure \ref{fig:wage_differential} plots the median of the last non-municipal sector wage by partisanship status. The gap between partisan and non-partisan wages is the highest for the beginning of the mayoral term, which is often associated with more intense patronage activity.

@@

Additionally, these same workers tend to have a longer experience working in the formal sector prior to employment. Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts. They have experienced both longer bouts of employment in the formal sector, as well as higher compensations in their tenure. While it is clear that these hires are a reward for previous contributions, it is not clear whether a party member who is hired into the bureaucracy is more likely to donate to their party. In order to determine this, it would be necessary to compare party members just before and after joining the local public sector, and see whether or not they are more likely to give campaign contributions.

@@

In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from wealthier strata of the population. Additionally, they have been working for longer time periods in the formal sector. However, on average, party members who enter the bureaucracy are less educated and older than their non-partisan counterparts, as evident in figure \ref{fig:partisan_edu_et_age}. Finally, while it is not clear what

\begin{figure}[H]
    \centering
    \includegraphics[width = 7cm, height = 8cm]{figures/partisanship/plot_edu_mean.pdf}
    \includegraphics[width = 7cm, height = 8cm]{figures/partisanship/plot_age_mean.pdf}
    \caption{\textbf{Heterogeneity in Education and Age.}}
    \label{fig:partisan_edu_et_age}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. This visualization includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}
